---
title: "Machine Learning with R"
author: "Kevins_Kacha"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r setup1, echo=TRUE}
loan_data <- readRDS("D:/kacha R workspace/loan_data.rds")
head(loan_data, 10)
library(caret)
library(data.tree)
library(caTools)
library(rpart)
library(rpart.plot)
library(tidyverse)
names(loan_data)
```

Summary of default rates accordindg to loan purpose.
```{r setup2, echo=TRUE}
default_rates <- loan_data |> 
  group_by(loan_purpose) |>      
  summarise(n_customers = n(),
            customers_default = sum(loan_default == "yes"),
            default_percent = 100 * mean(loan_default == "yes"))
default_rates
```

A visual presentation of the default rate considering loan purpose.
```{r setup3, echo=TRUE}
ggplot(data = default_rates, mapping = aes(x = loan_purpose, y = default_percent)) +
  geom_bar(stat = "identity", fill = "#006EA1", color = "white") +
           labs(title = "Loan Default Rate by Purpose of Loan",
                x = "Loan Purpose",
                y = "Default Percentage") +
             theme_light()
```

Summarizing default rate considering application type(Individual or group)

```{r setup4, echo=TRUE}
application_typ<-loan_data |> 
  group_by(application_type) |> 
  summarise(n_customers=n(),
            customers_default = sum(loan_default == "yes"),
            default_percent = 100 * mean(loan_default == "yes"))

application_typ
```

Visual presentation of the above.
```{r setup5, echo=TRUE}
ggplot(data = application_typ, mapping = aes(x = application_type, y = default_percent)) +
  geom_bar(stat = "identity", fill = "#006EA1", color = "white") +
  labs(title = "Loan Default Rate by application type",
       x = "application type",
       y = "Default Percentage") +
  theme_light()
```

Is there correlation between the loan given and the interest rate?

```{r setup7, echo=TRUE}
cor(loan_data$loan_amount, loan_data$interest_rate)
```


```{r setup8, echo=TRUE}
# homeownership
home<-loan_data |> 
  group_by(homeownership) |> 
  summarise(n_customers=n(),
            customers_default = sum(loan_default == "yes"),
            default_percent = 100 * mean(loan_default == "yes"))
home
```


```{r setup9, echo=TRUE}
ggplot(data = home, mapping = aes(x = homeownership, y = default_percent)) +
  geom_bar(stat = "identity", fill = "#006EA1", color = "white") +
  labs(title = "Loan Default Rate by homeownership",
       x = "homeownership",
       y = "Default Percentage") +
  theme_light()
```


```{r setup10, echo=TRUE}
# based on the term of payment
term_payment<-loan_data |> 
  group_by(term) |> 
  summarise(n_customers=n(),
            customers_default = sum(loan_default == "yes"),
            default_percent = 100 * mean(loan_default == "yes"))
term_payment
```


```{r setup11, echo=TRUE}
ggplot(data = term_payment, mapping = aes(x = term, y = default_percent)) +
  geom_bar(stat = "identity", fill = "#006EA1", color = "white") +
  labs(title = "Loan Default Rate by term_payment",
       x = "term_payment",
       y = "Default Percentage") +
  theme_light()
```

Based on years credit history

```{r setup12, echo=TRUE}
YearsCreditWorth<-loan_data |> 
  group_by(years_credit_history) |> 
  summarise(n_customers=n(),
            customers_default = sum(loan_default == "yes"),
            default_percent = 100 * mean(loan_default == "yes"))|> 
  arrange(desc(n_customers)) |> 
  head(19) |> 
  arrange(desc(default_percent)) |> 
  head(5)
YearsCreditWorth
```


```{r setup13, echo=TRUE}
Bankruptcy<-loan_data |> 
  group_by(history_bankruptcy) |> 
  summarise(n_customers=n(),
            customers_default = sum(loan_default == "yes"),
            default_percent = 100 * mean(loan_default == "yes"))
Bankruptcy
```


```{r setup14, echo=TRUE}
ggplot(data =Bankruptcy , mapping = aes(x = history_bankruptcy, y = default_percent)) +
  geom_bar(stat = "identity", fill = "#006EA1", color = "white") +
  labs(title = "Loan Default Rate by Bankruptcy",
       x = "Bankruptcy",
       y = "Default Percentage") +
  theme_light()
```

splitting the data  into training and testing data

```{r setup16, echo=TRUE}
dim(loan_data)


sample<- sample.split(loan_data$loan_default, SplitRatio = 0.8)
train<-subset(loan_data, sample == TRUE)
dim(train)
test<-subset(loan_data, sample==FALSE)
dim(test)
```

tree classifier
```{r setup17, echo=TRUE}
tree<-rpart(loan_default~., data = train)
tree
```

prediction using the test

```{r setup18, echo=TRUE}
tree.Default.predicted<-predict(tree, test, type = "class")
```

confusionMatrix for evaluating the model

```{r setup19, echo=TRUE}
confusionMatrix(tree.Default.predicted, test$loan_default)
```
rpartplot

```{r setup25, echo=TRUE}
prp(tree)
```


```{r setup20, echo=TRUE}
by3<-head(loan_data, 4)
by3.pred<-predict(tree, by3, type= "class")
by3.pred
```


