---
title: "Classification of iris using logistic regression"

author: "Kevins_Kacha"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Relevant packages

Calling for the relevant libraries that will aid in our task.

Rpart helps in checking for the relationship that exist between the classes.

Rpart.plot aids in drawing the decision tree.

We wish to classify the species iris data based on the flower attributes, including the sepal.length, sepal.width, petal.length and petal.width using the decision tree or logistic regression. 

```{r setup1, echo=TRUE}
library(rpart)
library(rpart.plot)
data("iris")
library(caret)
library(e1071)
```

## Data manipulation

Based on our dataset iris the data is classified based in the species ie setosa , virginica and versicolor as below. 
```{r setup18, echo=TRUE}
head(iris)
```
## Randomisation of the data

Our aim is to mix the data up before subsettung the train data and the testing data.

We assign randomly generated numbers which are uniformly distributed and arrange them in asceding order to mix up the dataset.

According to the glimple below, the data is now mixed up.
```{r setup2, echo=TRUE}
set.seed(500)
g<-runif(nrow(iris))
iris_ran<-iris[order(g),]
head(iris_ran)
```
## Fitting the model 
We proceed and select first 100 rows as training data and fit a model on it using rpart using classification method.
```{r setup3, echo=TRUE}
model1<-rpart(Species~., data =iris_ran[1:100,], method = "class")
model1
```
# The decision tree 
The decision tree gives a clear picture of the classification based on the features evident in the model

```{r setup4, echo=TRUE}

rpart.plot(model1, type = 4, fallen.leaves = T, extra = 104 )
```
According to our plot , it is observed that the setosa species had petal.length less than 2.5, versicolor and virginica had petal.length >= 2.5. They only differ in petal width as versicolor is <1.8 and virginica is >= 1.8.

The sepal.length and sepal.width does not influence the classification of species.
## Testing the model

We tested the model on the remaining 50 rows to evaluate the goodness of fit.
```{r setup5, echo=TRUE}
model.predict<-predict(model1,iris_ran[101:150,], type = "class")
model.predict
```
## Prediction accuracy

The function confusionmatrix in caTools helps to check the level of prediction accuracy.

Based on our model, the prediction accuracy is 98%. It predicted all setosa species correctly, out of 15 versicolor it predicted 14 of them correctly and lastly it predicted all the virginica species correctly.
```{r setup6, echo=TRUE}
confusionMatrix(iris_ran[101:150,5], reference = model.predict)
```

